### 2019-04-03 Thesis_Latex/ThesisFinal/Chapter5/Chapter5.tex
update final thesis

\chapter{Lane Following}
In this chapter we present a Lane Following alghorithm based on Adaptive Model Predictive Control. First we introduced the case-study model, then we designed the controller and the overall procedure scheme. Finally we verified the proposed strategy in different scenarios.
\section{Problem Formulation}
A lane-following system is a control system that keeps the vehicle traveling along the centerline of a highway lane, while maintaining a user-set velocity. Figure \ref{fig:laneFollowing} illustrates a typical lane following scenario.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.75\textwidth]{../figure/laneFollowing/laneFollowing.pdf}
	\caption{Problem description of a lane following system.}
	\label{fig:laneFollowing}
\end{figure}

@@

The vehicle is modelled at each axis through a bicycle model in which the two wheels are lumped into a single wheel. It will be assumed that a simple linear tire model can be used due to the considered speeds. It is also assumed that the small angles approximation can be used.
The equations describing the yaw and lateral motion are:
\begin{equation}
	\label{eqn:yaw_lateral_motion}
	\begin{array}{ll}
	m\dot{V_y}=F_{yf}+F_{yr}\\
	I_Z\ddot{\psi}=l_FF_{yf}-l_RF_{yr}
	\end{array}
\end{equation}
where $F_{yf}$ and $F_{yr}$ are the tire forces at the front and rear axles of the vehicle, which are a function of the cornering stifness of the tires $C_F$ and $C_R$, the steering angle $\delta$ and the vehicle states.
The ATLASCAR2 travels with a longitudinal velocity $V_x$ and a lateral velocity $V_y$. These two velocities can be used to form a vector describing the resultant velocity and its direction. In this case it is also possible to find the slip angles, the difference between the actual travelling direction of the and where it is pointed (in the previous chapter we assumed that the ATLASCAR2 does not slip, so any slippage is thus considered as an external disturbance):
\begin{equation}
	\label{eqn:alpha_f_alpha_r}
	\begin{array}{ll}
	\alpha_f = \delta-\theta_{vf}\\
	\alpha_r = -\theta_{vr}
	\end{array}
\end{equation}

@@

\begin{equation}
\label{eqn:theta_vf}
\theta_{vf}=\text{atan}\Bigg(\frac{V_y+l_F\dot{\psi}}{V_x}\Bigg)
\end{equation}
\begin{equation}
\label{eqn:theta_vr}
\theta_{vr}=\text{atan}\Bigg(\frac{V_y-l_R\dot{\psi}}{V_x}\Bigg)
\end{equation}

@@

\begin{equation}
	\label{eqn:lateral_tire_forces}
	\begin{array}{ll}
	F_{yf} = 2C_F(\delta-\theta_{vf})\\
	F_{yr} = -2C_R\theta_{vr}
	\end{array}
\end{equation}

Combining (\ref{eqn:yaw_lateral_motion}) and (\ref{eqn:lateral_tire_forces}) gives the following vehicle lateral model from parameters: 
\begin{equation}
\label{eqn:lateral_dynamics_simple_model}
\begin{array}{ll}
\dot{\vec{x}}_{\text{lat}} =\vec{A}_g \vec{x}_{\text{lat}}+ \vec{B}_g \vec{u}_{\text{lat}}\\
\vec{y}_{\text{lat}} =\vec{C}_g \vec{x}_{\text{lat}} + \vec{D}_g \vec{u}_{\text{lat}}
\end{array}
\end{equation}
where the input is the steering angle in radians, and the outputs are the lateral velocity in meters per second and yaw angle rate in radians per second:
\begin{equation}
\vec{x}_{\text{lat}} = \begin{bmatrix}
V_y\\\dot{\psi}
\end{bmatrix}
\qquad
\vec{u}_{\text{lat}} = \delta
\end{equation}
while the matrices $\vec{A}_g$,$\vec{B}_g$,$\vec{C}_g$ and $\vec{D}_g$ are the following:
\begin{equation}
\begin{array}{cc}
\vec{A}_g=
\begin{bmatrix}
\displaystyle -\frac{2C_F+2C_R}{mV_x}&\displaystyle -\frac{2C_Fl_F-2C_Rl_R}{mV_x} - V_x\\
\displaystyle -\frac{2C_Fl_F-2C_Rl_R}{I_ZV_x}&\displaystyle -\frac{2C_Fl_F^2+2C_Rl_R^2}{I_ZV_x}
\end{bmatrix},
\\\\
\vec{B}_g=\begin{bmatrix}
2C_F/m\\2C_Fl_F/I_Z
\end{bmatrix},
\qquad
\vec{C}_g=\begin{bmatrix}
1&0\\0&1
\end{bmatrix}=
\vec{I}_2, 
\qquad
\vec{D}_g=\begin{bmatrix}
0\\0
\end{bmatrix}=
\vec{0}_{2\times1}.
\end{array}
\end{equation}
Recapping the parameters in the previous matrices are:
\begin{itemize}
	\item $V_x$ is the longitudinal velocity of the car;	
	\item $m$ is the total mass parameter; 
	\item $I_Z$ is the yaw moment of inertia parameter;
	\item $l_F$ and $l_R$ are the longitudinal distances from center of gravity to front and rear tires parameters;
	\item $C_F$ and $C_R$ are the cornering stiffnesses of front and rear tires parameters.
\end{itemize}
\subsection{Augmented Model for Lateral Dynamics}
The goal for the driver steering model is to keep the vehicle in its lane and follow the curved road by controlling the front steering angle. Denote by $e_1$ the offset to the center line and $e_2$, the relative angle to the center line. This goal is achieved by driving the yaw angle error $e_2 = \psi -\psi_{\text{des}}$ and lateral displacement error $e_1$ to zero ($\dot{e}_1 = V_xe_2+V_y$). We can incorporate these two paramenters in the augmented model:
\begin{equation}
\label{eqn:lateral_dynamics_augmented_model}
\begin{array}{ll}
\dot{\vec{x}}_{\text{aug}} =\vec{A}_a \vec{x}_{\text{aug}}+ \vec{B}_a \vec{u}_{\text{aug}}\\
\vec{y}_{\text{aug}} = \vec{C}_a \vec{x}_{\text{aug}} + \vec{D}_a \vec{u}_{\text{aug}}
\end{array}
\end{equation}
where the states and the inputs are:
\begin{equation}
\vec{x}_{\text{aug}} = \begin{bmatrix}
V_y\\\dot{\psi}\\e_1\\e_2
\end{bmatrix},
\qquad
\vec{u}_{\text{aug}} = 
\begin{bmatrix}
\delta\\\dot{\psi}_{\text{des}}
\end{bmatrix}
\end{equation}
while the matrices $\vec{A}_a$,$\vec{B}_a$,$\vec{C}_a$ and $\vec{D}_a$ are the following:
\begin{equation}
\begin{array}{cc} 
\vec{A}_a=\begin{bmatrix}
\vec{A}_g&\vec{0}_{2\times2}\\
\vec{I}_2&\begin{matrix}
0&V_x\\
0&0
\end{matrix}
\end{bmatrix},
\qquad
\vec{B}_a=\begin{bmatrix}
\vec{B}_g&\vec{0}_{2\times1}\\
0&0\\
0&-1
\end{bmatrix},
\\\\
\vec{C}_a=\begin{bmatrix}
\vec{0}_{2\times2}&\vec{I}_2
\end{bmatrix}, 
\qquad
\vec{D}_a=
\vec{0}_{2\times2}. 
\end{array}
\end{equation}

@@

\section{Design of Adaptive Model Predictive Control}
Before talking about the design of the controller, it is necessary to highlight which are all the elements that make up our control scheme. The overall framework for a lane keeping assist is depicted in the Figure \ref{fig:scheme_lane_following}.
%OVERALL SCHEME LANE FOLLOWING
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{../figure/lane_following_AMPC.pdf}
	\caption{Overall procedure scheme lane following.}
	\label{fig:scheme_lane_following}
\end{figure}

@@

%VEHICLE DYNAMICS
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{../figure/lane_following_AMPC_vehicle_dynamics.pdf}
	\caption{Vehicle Dynamics of the overall scheme lane following.}
	\label{fig:scheme_lane_following_vehicle_dynamics}
\end{figure}
We created an Adaptive MPC controller with a prediction model that has six states, three outputs (longitudinal velocity, lateral deviation, relative yaw angle), two manipulated signals (acceleration and steering) and one measured disturbance (desired yaw rate).

In order to design a valid lane keeping algorithm based on MPC we have set the constraints for manipulated variables and the scale factors. In particular the control
variables are constrained as follows:
\begin{equation}
\label{eqn:constraint_LKA}
\begin{array}{cc}
	\SI{-3}{m/s^2}\leq a\leq\SI{3}{m/s^2}\;,\\\\
	\SI{-1.13}{rad}\leq \delta \leq \SI{1.13}{rad}\;,
\end{array}
\end{equation}

Moreover we have specified the weights in the standard MPC cost function. The third output, yaw angle, is allowed to float because there are only two manipulated variables to make it a square system. In this controller, there is no steady-state error in the yaw angle as long as the second output, lateral deviation, reaches 0 at steady state. Finally we have also penalized acceleration change more for smooth driving experience. This controller uses a linear model for the vehicle dynamics and updates the model online as the longitudinal velocity varies.
The dynamics of the sensors (Figure \ref{fig:scheme_lane_following_sensor_dynamics}) allows us to calculate two fundamental parameters to reach the goal of keeping the vehicle in its lane and following the curved road by controlling the front steering angle:
\begin{itemize}
	\item $e_1$ which is the offset to the center line,
	\item $e_2$ which is the relative angle to the center line.
\end{itemize}
where we want to drive the yaw angle error and the lateral displacement error to zero to achieve our goal. The product of the road curvature and the longitudinal velocity is modeled as a measured disturbance.
%SENSOR DYNAMICS
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{../figure/lane_following_AMPC_sensor_dynamics.pdf}
	\caption{Sensor Dynamics of the overall scheme lane following.}
	\label{fig:scheme_lane_following_sensor_dynamics}
\end{figure}

\section{Simulation Results}
The proposed adaptive MPC algorithm is designed in the MATLAB/Simulink and validated through simulations considering different paths. The objective of these tests is to evaluate the behavior of the proposed control strategy in critical situations. The parameters used in the lane following simulations are summarized in table \ref{tab:params}:
\begin{table}[!h]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		\bfseries Parameters          & \bfseries Values   & \bfseries Description   \\
		\hline
		$m$          & \SI{1575}{kg}  & Mass of the vehicle            \\
		$I_z$         & \SI{2875}{kgm^2}   & Inertia Moment            \\
		$l_F$           & \SI{1.2}{m}     & Distance from COG to front axle         \\
		$l_R$         & \SI{1.6}{m}    & Distance from COG to rear axle           \\
		$C_F$          & \SI{19000}{\newton/rad}&Front cornering stiffness      \\
		$C_R$           & \SI{33000}{\newton/rad}&Rear cornering stiffness  \\
		$\tau$             & 0.2    & Time constant            \\
		$V_0$           & \SI{15}{m/s} & Initial Velocity      \\
		$V_{\text{set}}$          & \SI{20}{m/s} & Driver-set Velocity  \\
		$T_s$         & \SI{0.02}{s} & Sampling Time
		         \\
		\hline
	\end{tabular}
	\caption{Parameters of Vehicle Dynamics and Road Curvature.}
	\label{tab:params}
\end{table}

@@

	\caption{Desired sinusoidal path and its curvature of the ATLASCAR2 in a simulation of \SI{20}{s} with initial position in $(0,0)$.}

@@

In practice this type of route can be followed by the vehicle when it has to make a slalom between cones or it must go up some hairpin bends in the mountains. Moreover the following figures show the trend of the main parameters confirming that the control strategy used allows the vehicle to follow the path. In particular we simulated also a small error in the sensor dynamics in order to make the simulation more realistic: we added a 3 percent error to the longitudinal velocity and this is evident from the small noise in the graphs of the steering angle (Figure \ref{fig:steering_laneFollowing}) and the lateral deviation (Figure \ref{fig:lateral_deviation_laneFollowing}).
Figure {\ref{fig:longitudinal_velocity_laneFollowing}} shows the evolution of the vehicle longitudinal velocity. Also in this simulation the velocity is equal to the initial condition for longitudinal velocity parameter $V_0$. At run time, we can note that $V_x$ reaches the predefined value of \SI{20}{m/s} and then it stabilizes near the cruising speed because it continues to vary the steering angle to adapt to the path to be followed. In this scenario the lateral deviation and the relative angle yaw do not converge to zero but oscillate. We can however assert that this type of oscillation is due to the type of path we are following. Furthermore the limits of this oscillation are very small indeed: in particular the lateral deviation is less than \SI{1.5}{cm} while the maximum relative yaw angle is \SI{0.04}{rad}. We can therefore state that even in this case the vehicle correctly follows the lane.
%LANE FOLLOWING SINUSOIDAL PATH - SIGNALS
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LongitudinalVelocityVsTime.pdf}
		\label{fig:longitudinal_velocity_laneFollowing}
			\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralVelocityVsTime.pdf}
			\label{fig:lateral_velocity_laneFollowing}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/AccelerationVsTime.pdf}
		\label{fig:acceleration_laneFollowing}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/SteeringAngleVsTime.pdf}
		\label{fig:steering_laneFollowing}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/RelativeYawAngleVsTime.pdf}
		\label{fig:relative_yaw_angle_laneFollowing}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralDeviationVsTime.pdf}
		\label{fig:lateral_deviation_laneFollowing}
	\caption{Time signals of the ATLASCAR2 in the simulation with a sinusoidal path.}
	\label{fig:laneFollowing_signals}
\subsection{Circular/Elliptic Path}
In the last simulation we considered that the vehicle must follow an elliptic path with a semi-major axis of \SI{30}{m} and semi-minor axis of \SI{15}{m} depicted in Figure \ref{fig:reference_laneFollowing_circular}. Also in this case the simulation time is \SI{20}{s}.
To describe the elliptic path and to evaluate its curvature, depicted in Figure \ref{fig:curvature_laneFollowing_circular}, the equations are as follows:
	X_\text{ref}&=\displaystyle 15\cos\Bigg(\frac{V_x\cdot t}{60}\Bigg)\\
	Y_\text{ref}&=\displaystyle 30\sin\Bigg(\frac{V_x\cdot t}{60}\Bigg)  
	\end{aligned} \quad\text{with}\quad t\in[0,20]\SI{}{s}
% ELLIPTIC PATH
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Reference_circular.pdf}
		\label{fig:reference_laneFollowing_circular}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Curvature_circular.pdf}
		\label{fig:curvature_laneFollowing_circular}
	\caption{Desired elliptic path and its curvature of the ATLASCAR2 in a simulation of \SI{20}{s} with initial position in $(0,30)$.}
	\label{fig:laneFollowing_desired_circular}

>>>>

\chapter{Lane Following}
In this chapter we present a Lane Following alghorithm based on Adaptive Model Predictive Control. First we introduced the case-study model, then we designed the controller and the overall procedure scheme. Finally we verified the proposed strategy in different scenarios considering a double lane change, sinusoidal and elliptic/circular paths.
\section{Problem Formulation}
A lane-following system is a control system that keeps the vehicle traveling along the centerline of a highway lane, while maintaining a user-set velocity. Figure \ref{fig:laneFollowing} illustrates a typical lane following scenario.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.75\textwidth]{./figure/laneFollowing/laneFollowing.pdf}
	\caption{Problem description of a lane following system.}
	\label{fig:laneFollowing}
\end{figure}

@@

The vehicle is modelled at each axis through a bicycle model in which the two wheels are lumped into a single wheel. It will be assumed that a simple linear tire model can be used due to the considered speeds. It is also assumed that the small angles approximation can be used.
The equations describing the yaw and lateral motion are:
\begin{equation}
	\label{eqn:yaw_lateral_motion}
	\begin{array}{ll}
	m\dot{V_y}=F_{yf}+F_{yr}\\
	I_Z\ddot{\psi}=l_FF_{yf}-l_RF_{yr}
	\end{array}
\end{equation}
where $F_{yf}$ and $F_{yr}$ are the tire forces at the front and rear axles of the vehicle, which are a function of the cornering stifness of the tires $C_F$ and $C_R$, the steering angle $\delta$ and the vehicle states.
The ATLASCAR2 travels with a longitudinal velocity $V_x$ and a lateral velocity $V_y$. These two velocities can be used to form a vector describing the resultant velocity and its direction. In this case it is also possible to find the slip angles, the difference between the actual travelling direction of the and where it is pointed (in the previous chapter we assumed that the ATLASCAR2 does not slip, so any slippage is thus considered as an external disturbance):
\begin{equation}
	\label{eqn:alphas}
	\alpha_f = \delta-\theta_{vf},\qquad\quad
	\alpha_r = -\theta_{vr}
\end{equation}

@@

\begin{equation}
\label{eqn:theta_v}
\theta_{vf}=\text{atan}\Bigg(\frac{V_y+l_F\dot{\psi}}{V_x}\Bigg),
\qquad \quad
\theta_{vr}=\text{atan}\Bigg(\frac{V_y-l_R\dot{\psi}}{V_x}\Bigg)
\end{equation}

@@

\begin{equation}
	\label{eqn:lateral_tire_forces}
	F_{yf} = 2C_F(\delta-\theta_{vf}),\qquad\quad
	F_{yr} = -2C_R\theta_{vr}
\end{equation}

Combining (\ref{eqn:yaw_lateral_motion}) and (\ref{eqn:lateral_tire_forces}) gives the following vehicle lateral model from parameters: 
\begin{equation}
\label{eqn:lateral_dynamics_simple_model}
\begin{array}{ll}
\dot{\vec{x}}_{\text{lat}} =\vec{A}_g \vec{x}_{\text{lat}}+ \vec{B}_g \vec{u}_{\text{lat}}\\
\vec{y}_{\text{lat}} =\vec{C}_g \vec{x}_{\text{lat}} + \vec{D}_g \vec{u}_{\text{lat}}
\end{array}
\end{equation}
where the input is the steering angle in radians, and the outputs are the lateral velocity in meters per second and yaw angle rate in radians per second:
\begin{equation}
\vec{x}_{\text{lat}} = \begin{bmatrix}
V_y\\\dot{\psi}
\end{bmatrix}
\qquad
\vec{u}_{\text{lat}} = \delta
\end{equation}
while the matrices $\vec{A}_g$,$\vec{B}_g$,$\vec{C}_g$ and $\vec{D}_g$ are the following:
\begin{equation}
\begin{array}{cc}
\vec{A}_g=
\begin{bmatrix}
\displaystyle -\frac{2C_F+2C_R}{mV_x}&\displaystyle -\frac{2C_Fl_F-2C_Rl_R}{mV_x} - V_x\\
\displaystyle -\frac{2C_Fl_F-2C_Rl_R}{I_ZV_x}&\displaystyle -\frac{2C_Fl_F^2+2C_Rl_R^2}{I_ZV_x}
\end{bmatrix},
\\\\
\vec{B}_g=\begin{bmatrix}
2C_F/m\\2C_Fl_F/I_Z
\end{bmatrix},
\qquad
\vec{C}_g=\begin{bmatrix}
1&0\\0&1
\end{bmatrix}=
\vec{I}_2, 
\qquad
\vec{D}_g=\begin{bmatrix}
0\\0
\end{bmatrix}=
\vec{0}_{2\times1}.
\end{array}
\end{equation}
Recapping the parameters in the previous matrices are:
\begin{itemize}
	\item $V_x$ is the longitudinal velocity of the car;	
	\item $m$ is the total mass parameter; 
	\item $I_Z$ is the yaw moment of inertia parameter;
	\item $l_F$ is the longitudinal distances from center of gravity to front tire parameter;
	\item $l_R$ is the longitudinal distances from center of gravity to rear tire parameter;
	\item $C_F$ is the cornering stiffnesses of front tire parameter;
	\item $C_R$ is the cornering stiffnesses of rear tire parameter;
\end{itemize}
\subsection{Augmented Model for Lateral Dynamics}
The goal for the driver steering model is to keep the vehicle in its lane and follow the curved road by controlling the front steering angle. Denote by $e_1$ the offset to the center line and $e_2$, the relative angle to the center line. This goal is achieved by driving the yaw angle error $e_2 = \psi -\psi_{\text{des}}$ and lateral displacement error $e_1$ to zero ($\dot{e}_1 = V_xe_2+V_y$). We can incorporate these two paramenters in the augmented model:
\begin{equation}
\label{eqn:lateral_dynamics_augmented_model}
\begin{array}{ll}
\dot{\vec{x}}_{\text{aug}} =\vec{A}_a \vec{x}_{\text{aug}}+ \vec{B}_a \vec{u}_{\text{aug}}\\
\vec{y}_{\text{aug}} = \vec{C}_a \vec{x}_{\text{aug}} + \vec{D}_a \vec{u}_{\text{aug}}
\end{array}
\end{equation}
where the states and the inputs are:
\begin{equation}
\vec{x}_{\text{aug}} = \begin{bmatrix}
V_y\\\dot{\psi}\\e_1\\e_2
\end{bmatrix},
\qquad
\vec{u}_{\text{aug}} = 
\begin{bmatrix}
\delta\\\dot{\psi}_{\text{des}}
\end{bmatrix}
\end{equation}
while the matrices $\vec{A}_a$,$\vec{B}_a$,$\vec{C}_a$ and $\vec{D}_a$ are the following:
\begin{equation}
\begin{array}{cc} 
\vec{A}_a=\begin{bmatrix}
\vec{A}_g&\vec{0}_{2\times2}\\
\vec{I}_2&\begin{matrix}
0&V_x\\
0&0
\end{matrix}
\end{bmatrix},
\qquad
\vec{B}_a=\begin{bmatrix}
\vec{B}_g&\vec{0}_{2\times1}\\
0&0\\
0&-1
\end{bmatrix},
\\\\
\vec{C}_a=\begin{bmatrix}
\vec{0}_{2\times2}&\vec{I}_2
\end{bmatrix}, 
\qquad
\vec{D}_a=
\vec{0}_{2\times2}. 
\end{array}
\end{equation}

@@

\section{Design of Adaptive Model Predictive Control}
Before talking about the design of the controller, it is necessary to highlight which are all the elements that make up our control scheme. The overall framework for a lane keeping assist is depicted in the Figure \ref{fig:scheme_lane_following}.
%OVERALL SCHEME LANE FOLLOWING
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{./figure/lane_following_AMPC.pdf}
	\caption{Overall procedure scheme lane following.}
	\label{fig:scheme_lane_following}
\end{figure}

@@

%VEHICLE DYNAMICS
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{./figure/lane_following_AMPC_vehicle_dynamics.pdf}
	\caption{Vehicle Dynamics of the overall scheme lane following.}
	\label{fig:scheme_lane_following_vehicle_dynamics}
\end{figure}
We created an Adaptive MPC controller with a prediction model that has six states, three outputs (longitudinal velocity, lateral deviation, relative yaw angle), two manipulated signals (acceleration and steering) and one measured disturbance (desired yaw rate).

In order to design a valid lane keeping algorithm based on MPC we have set the constraints for manipulated variables and the scale factors. In particular the control
variables are constrained as follows:
\begin{equation}
\label{eqn:constraint_LKA}
\begin{array}{cc}
	\SI{-3}{m/s^2}\leq a\leq\SI{3}{m/s^2}\qquad\quad
	\SI{-1.13}{rad}\leq \delta \leq \SI{1.13}{rad}\;,
\end{array}
\end{equation}

Moreover we have specified the weights in the standard MPC cost function. The third output, yaw angle, is allowed to float because there are only two manipulated variables to make it a square system. In this controller, there is no steady-state error in the yaw angle as long as the second output, lateral deviation, reaches 0 at steady state. Finally we have also penalized acceleration change more for smooth driving experience. This controller uses a linear model for the vehicle dynamics and updates the model online as the longitudinal velocity varies.
The dynamics of the sensors (Figure \ref{fig:scheme_lane_following_sensor_dynamics}) allows us to calculate two fundamental parameters to reach the goal of keeping the vehicle in its lane and following the curved road by controlling the front steering angle:
\begin{itemize}
	\item $e_1$ which is the offset to the center line,
	\item $e_2$ which is the relative angle to the center line.
\end{itemize}
where we want to drive the yaw angle error and the lateral displacement error to zero to achieve our goal. The product of the road curvature and the longitudinal velocity is modeled as a measured disturbance.
%SENSOR DYNAMICS
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{./figure/lane_following_AMPC_sensor_dynamics.pdf}
	\caption{Sensor Dynamics of the overall scheme lane following.}
	\label{fig:scheme_lane_following_sensor_dynamics}
\end{figure}

\section{Simulation Results}
The proposed adaptive MPC algorithm is designed in the MATLAB/Simulink and validated through simulations considering different paths. The objective of these tests is to evaluate the behavior of the proposed control strategy in critical situations. The parameters used in the lane following simulations are summarized in table \ref{tab:params}:
\begin{table}[!h]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		\bfseries Parameters          & \bfseries Values   & \bfseries Description   \\
		\hline
		$m$          & \SI{1575}{kg}  & Mass of the vehicle            \\
		$I_z$         & \SI{2875}{kgm^2}   & Inertia Moment            \\
		$l_F$           & \SI{1.2}{m}     & Distance from COG to front axle         \\
		$l_R$         & \SI{1.6}{m}    & Distance from COG to rear axle           \\
		$C_F$          & \SI{19000}{\newton/rad}&Front cornering stiffness      \\
		$C_R$           & \SI{33000}{\newton/rad}&Rear cornering stiffness  \\
		$\tau$             & \SI{0.2}    & Time constant            \\
		$V_0$           & \SI{15}{m/s} & Initial Velocity      \\
		$V_{\text{set}}$          & \SI{20}{m/s} & Driver-set Velocity  \\
		$T_s$         & \SI{0.02}{s} & Sampling Time
		         \\
		\hline
	\end{tabular}
	\caption{Parameters of Vehicle Dynamics and Road Curvature.}
	\label{tab:params}
\end{table}

@@

	\caption{Desired sinusoidal path and its curvature of the ATLASCAR2 in a simulation of \SI{20}{s}.}

@@

In practice this type of route can be followed by the vehicle when it has to make a slalom between cones or it must go up some hairpin bends in the mountains. Moreover the following figures show the trend of the main parameters confirming that the control strategy used allows the vehicle to follow the path. In particular we simulated also a small error in the sensor dynamics in order to make the simulation more realistic: we added a 3 percent error to the longitudinal velocity and this is evident from the small noise in the graphs of the steering angle (Figure \ref{fig:steering_laneFollowing}) and the lateral deviation (Figure \ref{fig:lateral_deviation_laneFollowing}).
Figure {\ref{fig:longitudinal_velocity_laneFollowing}} shows the evolution of the vehicle longitudinal velocity. Also in this simulation the velocity is equal to the initial condition for longitudinal velocity parameter $V_0$. At run time, we can note that $V_x$ reaches the predefined value of \SI{20}{m/s} and then it stabilizes near the cruising speed because it continues to vary the steering angle to adapt to the path to be followed. In this scenario the lateral deviation and the relative angle yaw do not converge to zero but oscillate. We can however assert that this type of oscillation is due to the type of path we are following. Furthermore the limits of this oscillation are very small indeed: in particular the lateral deviation is less than \SI{1.5}{cm} while the maximum relative yaw angle is \SI{0.04}{rad}. We can therefore state that even in this case the vehicle correctly follows the lane.
%LANE FOLLOWING SINUSOIDAL PATH - SIGNALS
\begin{figure}[!h]
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LongitudinalVelocityVsTime.pdf}
		\subcaption{Longitudinal velocity $V_x$ w.r.t. time.}
		\label{fig:longitudinal_velocity_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
			\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralVelocityVsTime.pdf}
			\subcaption{Lateral velocity $V_y$ w.r.t. time.}
			\label{fig:lateral_velocity_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/AccelerationVsTime.pdf}
		\subcaption{Acceleration $u_1$ w.r.t. time.}
		\label{fig:acceleration_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/SteeringAngleVsTime.pdf}
		\subcaption{Steering angle $u_2$ w.r.t. time.}
		\label{fig:steering_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/RelativeYawAngleVsTime.pdf}
		\subcaption{Relative yaw angle $e_2$ w.r.t. time.}
		\label{fig:relative_yaw_angle_laneFollowing}
	\end{minipage}
	\begin{minipage}[t]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following/figure/LateralDeviationVsTime.pdf}
		\subcaption{Lateral deviation $e_1$ w.r.t. time.}
		\label{fig:lateral_deviation_laneFollowing}
	\end{minipage}
	\caption{Time signals of the ATLASCAR2 in the simulation with a sinusoidal path.}
	\label{fig:laneFollowing_signals}
\end{figure}

\subsection{Circular/Elliptic Path}
In the last simulation we considered that the vehicle must follow an elliptic path with a semi-major axis of \SI{30}{m} and semi-minor axis of \SI{15}{m} depicted in Figure \ref{fig:reference_laneFollowing_circular}. Also in this case the simulation time is \SI{20}{s}.
To describe the elliptic path and to evaluate its curvature, depicted in Figure \ref{fig:curvature_laneFollowing_circular}, the equations are as follows:
\begin{equation}
	X_\text{ref}=\displaystyle 15\cos\Bigg(\frac{V_x\cdot t}{60}\Bigg) \qquad
	Y_\text{ref}=\displaystyle 30\sin\Bigg(\frac{V_x\cdot t}{60}\Bigg)  
 \quad\text{with}\quad t\in[0,20]\SI{}{s}
\end{equation}
% ELLIPTIC PATH
\begin{figure}[!h]
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Reference_circular.pdf}
		\subcaption{}
		\label{fig:reference_laneFollowing_circular}
	\end{minipage}
	\begin{minipage}[t]{0.5\textwidth}
		\includegraphics[width=\textwidth]{../../MATLAB/lane_following_circular_path/figure/Curvature_circular.pdf}
		\subcaption{}
		\label{fig:curvature_laneFollowing_circular}
	\end{minipage}
	\caption{Desired elliptic path and its curvature of the ATLASCAR2 in a simulation of \SI{20}{s} with initial position in $(0,30)$.}
	\label{fig:laneFollowing_desired_circular}
\end{figure}
